---
title: "Regression Tree"
output: html_document
---

################### Global Model ###################

```{r}
library(readxl)
library(mlr)
library(rpart)
library(rpart.plot)
library(dplyr)
library(Hmisc)
library(ggplot2)
```
################### Reading the dataset ###################
```{r}
my_data <- read_excel('FinalModifiedData.xlsx')
my_data <- as.data.frame(my_data)

glimpse(my_data)

my_data$Station.Code<-NULL
my_data$Hour<-NULL
my_data$Year<-NULL
my_data$Month<-NULL
my_data$Day<-NULL

str(my_data)

```
################## Univariate Analysis ##################
```{r}

describe(my_data)

```
################## Histograms for distribution of data ##################
```{r}

ggplot(my_data, aes(x=my_data$Min.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data, aes(x=my_data$Max.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data, aes(x=my_data$Temp.Dry.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data, aes(x=my_data$Temp.Wet.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data, aes(x=my_data$Inst.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data, aes(x=my_data$Av.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data, aes(x=my_data$Relative.Humidity)) + geom_histogram()

```
```{r}
ggplot(my_data, aes(x=my_data$Temp...Pan.Water)) + geom_histogram()

```
```{r}
ggplot(my_data, aes(x=my_data$Pan.Evaporation)) + geom_histogram()

```
################## Box plots for outliers ##################
```{r}
# Create separate boxplots for each attribute
par(mfrow=c(1,3))
for(i in 1:3) {
	boxplot(my_data[,i], main=names(my_data)[i])
}

```
```{r}
par(mfrow=c(1,3))
for(i in 4:6) {
	boxplot(my_data[,i], main=names(my_data)[i])
}
```
```{r}
par(mfrow=c(1,3))
for(i in 8:9)
{
    
  boxplot(my_data[,i], main=names(my_data)[i])
}
```

```{r}
# change categorical values to numerical values

my_data[my_data=="C"]<-0
my_data[my_data=="E"]<-9
my_data[my_data=="S"]<-18
my_data[my_data=="W"]<-27
my_data[my_data=="N"]<-36
my_data[my_data=="NNE"]<-2
my_data[my_data=="ESE"]<-11
my_data[my_data=="SSW"]<- 20
my_data[my_data=="WNW"]<- 29
my_data[my_data=="V"]<- 99
my_data[my_data=="NE"]<- 5
my_data[my_data=="SE"]<- 14
my_data[my_data=="SW"]<- 23
my_data[my_data=="NW"]<- 32
my_data[my_data=="ENE"]<- 7
my_data[my_data=="SSE"]<- 16
my_data[my_data=="WSW"]<- 25
my_data[my_data=="NNW"]<- 34

my_data$Wind.direction <- as.numeric(as.character(my_data$Wind.direction))

str(my_data)

```
################### Visualizing the Correlation matrix ###################
```{r}
source("http://www.sthda.com/upload/rquery_cormat.r")
require("corrplot")
rquery.cormat(my_data)
```

```{r}

rmse_reg <- function(model_obj, testing = NULL, target = NULL) {
  #Calculates rmse for a regression decision tree
  #Arguments:
    # testing - test data set
    # target  - target variable (length 1 character vector)
    yhat <- predict(model_obj, newdata = testing)
  actual <- testing[[target]]
  sqrt(mean((yhat-actual)^2))
}

```
#############Creating the Task #############
```{r}
#Splitting the dataset

sample <- sample.int(n = nrow(my_data), size = floor(.75*nrow(my_data)), replace = F)
train <- my_data[sample, ]
test  <- my_data[-sample, ]

#Regression Tree:Task Creation
regr.task.train<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(train), 
                        target = "Max.Temperature")

regr.task.test<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(test), 
                        target = "Max.Temperature")

#Repeated cross validation
cv.folds <- makeResampleDesc("CV", iters = 3)

# Define model tuning algorithm ~ Random tune algorithm
random.tune <- makeTuneControlRandom(maxit = 5)

```
#############Learner for the model#############
```{r}

set.seed(1)

#Defining learner model
lrn <- makeLearner("regr.rpart") 

#Get parameter set of the learner
getParamSet(lrn)                        

# Define parameters of model  
model.params <- makeParamSet( 
  makeNumericParam("cp", lower = 0, upper = 0.1),
  makeIntegerParam("minsplit", lower = 5, upper = 20),
  makeIntegerParam("minbucket", lower = 2, upper = 10)
)

#Hyperparameter tuning
tuned.model <- tuneParams(learner    = lrn,
                             task       = regr.task.train,
                             resampling = cv.folds,
                             par.set    = model.params,
                             control    = random.tune,
                             measures = list(rmse,mae),
                             show.info  = TRUE)
tuned.model$x
```
############# Evaluation of Model #############
```{r}

lrn <- makeLearner("regr.rpart",par.vals = tuned.model$x)

res <- mlr::train(lrn,regr.task.train)
rpc = rpart.control(minsplit = tuned.model$x$minsplit, minbucket = tuned.model$x$minbucket, cp = tuned.model$x$cp
                    )
mod <- rpart(Max.Temperature ~ ., data = train, control = rpc)

#Calucating the RMSE of training set 
rmse_reg(mod, testing = train, target = "Max.Temperature")
getLearnerModel(res)

rpart.plot(res$learner.model,roundint=FALSE)

```
```{r}

(mlrpred   <- predict(res, newdata = test))

#Calcuating the RMSE of test set
performance(mlrpred, measures = rmse) 

```

################# LOCAL ####################
############## Station code: Amli ##############

```{r}

my_data1 <- read_excel('SCAmli.xlsx')
my_data1 <- as.data.frame(my_data1)

glimpse(my_data1)

my_data1$Station.Code<-NULL
my_data1$Hour<-NULL
my_data1$Year<-NULL
my_data1$Month<-NULL
my_data1$Day<-NULL

str(my_data1)

```
################## Histograms distribution of data ##################
```{r}
ggplot(my_data1, aes(x=my_data1$Min.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data1, aes(x=my_data1$Max.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data1, aes(x=my_data1$Temp.Dry.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data1, aes(x=my_data1$Temp.Wet.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data1, aes(x=my_data1$Inst.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data1, aes(x=my_data1$Av.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data1, aes(x=my_data1$Relative.Humidity)) + geom_histogram()

```
```{r}
ggplot(my_data1, aes(x=my_data1$Temp...Pan.Water)) + geom_histogram()

```
```{r}
ggplot(my_data1, aes(x=my_data1$Pan.Evaporation)) + geom_histogram()

```

```{r}
# change categorical values to numerical values

my_data1[my_data1=="C"]<-0
my_data1[my_data1=="E"]<-9
my_data1[my_data1=="S"]<-18
my_data1[my_data1=="W"]<-27
my_data1[my_data1=="N"]<-36
my_data1[my_data1=="NNE"]<-2
my_data1[my_data1=="ESE"]<-11
my_data1[my_data1=="SSW"]<- 20
my_data1[my_data1=="WNW"]<- 29
my_data1[my_data1=="V"]<- 99
my_data1[my_data1=="NE"]<- 5
my_data1[my_data1=="SE"]<- 14
my_data1[my_data1=="SW"]<- 23
my_data1[my_data1=="NW"]<- 32
my_data1[my_data1=="ENE"]<- 7
my_data1[my_data1=="SSE"]<- 16
my_data1[my_data1=="WSW"]<- 25
my_data1[my_data1=="NNW"]<- 34

my_data1$Wind.direction <- as.numeric(as.character(my_data1$Wind.direction))

str(my_data1)
```
################ Visualizing the Correlation matrix ################
```{r}
source("http://www.sthda.com/upload/rquery_cormat.r")
require("corrplot")
rquery.cormat(my_data1)
```

```{r}
#Splitting the dataset

sample1 <- sample.int(n = nrow(my_data1), size = floor(.75*nrow(my_data1)), replace = F)
train1 <- my_data1[sample1, ]
test1  <- my_data1[-sample1, ]

#Regression Tree:Task Creation
regr.task.train1<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(train1), 
                        target = "Max.Temperature")

regr.task.test1<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(test1), 
                        target = "Max.Temperature")

#Repeated cross validation
cv.folds <- makeResampleDesc("CV", iters = 3)

# Define model tuning algorithm ~ Random tune algorithm
random.tune <- makeTuneControlRandom(maxit = 5)
```

```{r}
set.seed(1)

#Defining learner model
lrn1 <- makeLearner("regr.rpart") 

#Get parameter set of the learner
getParamSet(lrn1)                        

# Define parameters of model  
model.params <- makeParamSet( 
  makeNumericParam("cp", lower = 0, upper = 0.1),
  makeIntegerParam("minsplit", lower = 5, upper = 20),
  makeIntegerParam("minbucket", lower = 2, upper = 10)
)

#Hyperparameter tuning
tuned.model1 <- tuneParams(learner    = lrn1,
                             task       = regr.task.train1,
                             resampling = cv.folds,
                             par.set    = model.params,
                             control    = random.tune,
                             measures = list(rmse,mae),
                             show.info  = TRUE)
tuned.model1$x
```

```{r}
lrn1 <- makeLearner("regr.rpart",par.vals = tuned.model1$x)

res1 <- mlr::train(lrn1,regr.task.train1)
rpc1 = rpart.control(minsplit = tuned.model1$x$minsplit, minbucket = tuned.model1$x$minbucket, cp = tuned.model1$x$cp
                    )
mod1 <- rpart(Max.Temperature ~ ., data = train1, control = rpc1)

#Calucating the RMSE of training set 
rmse_reg(mod1, testing = train1, target = "Max.Temperature")
getLearnerModel(res1)

rpart.plot(res1$learner.model,roundint=FALSE)

```

```{r}

(mlrpred1   <- predict(res1, newdata = test1))

#Calcuating the RMSE of test set

sc1 <- performance(mlrpred1, measures = rmse) 

print(sc1)

```

############### Station code: Chopadvav ###############

```{r}

my_data2 <- read_excel('SCChop.xlsx')
my_data2 <- as.data.frame(my_data2)

glimpse(my_data2)

my_data2$Station.Code<-NULL
my_data2$Hour<-NULL
my_data2$Year<-NULL
my_data2$Month<-NULL
my_data2$Day<-NULL

str(my_data2)

```

```{r}
# change categorical values to numerical values

my_data2[my_data2=="C"]<-0
my_data2[my_data2=="E"]<-9
my_data2[my_data2=="S"]<-18
my_data2[my_data2=="W"]<-27
my_data2[my_data2=="N"]<-36
my_data2[my_data2=="NNE"]<-2
my_data2[my_data2=="ESE"]<-11
my_data2[my_data2=="SSW"]<- 20
my_data2[my_data2=="WNW"]<- 29
my_data2[my_data2=="V"]<- 99
my_data2[my_data2=="NE"]<- 5
my_data2[my_data2=="SE"]<- 14
my_data2[my_data2=="SW"]<- 23
my_data2[my_data2=="NW"]<- 32
my_data2[my_data2=="ENE"]<- 7
my_data2[my_data2=="SSE"]<- 16
my_data2[my_data2=="WSW"]<- 25
my_data2[my_data2=="NNW"]<- 34

my_data2$Wind.direction <- as.numeric(as.character(my_data2$Wind.direction))

str(my_data2)
```
################## Histograms distribution of data ##################
```{r}
ggplot(my_data2, aes(x=my_data2$Min.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data2, aes(x=my_data2$Max.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data2, aes(x=my_data2$Temp.Dry.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data2, aes(x=my_data2$Temp.Wet.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data2, aes(x=my_data2$Inst.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data2, aes(x=my_data2$Av.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data2, aes(x=my_data2$Relative.Humidity)) + geom_histogram()

```
```{r}
ggplot(my_data2, aes(x=my_data2$Temp...Pan.Water)) + geom_histogram()

```
```{r}
ggplot(my_data2, aes(x=my_data2$Pan.Evaporation)) + geom_histogram()

```
################ Visualizing the Correlation matrix ################
```{r}
source("http://www.sthda.com/upload/rquery_cormat.r")
require("corrplot")
rquery.cormat(my_data2)
```

```{r}
#Splitting the dataset

sample2 <- sample.int(n = nrow(my_data2), size = floor(.75*nrow(my_data2)), replace = F)
train2 <- my_data2[sample2, ]
test2  <- my_data2[-sample2, ]

#Regression Tree:Task Creation
regr.task.train2<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(train2), 
                        target = "Max.Temperature")

regr.task.test2<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(test2), 
                        target = "Max.Temperature")

#Repeated cross validation
cv.folds <- makeResampleDesc("CV", iters = 3)

# Define model tuning algorithm ~ Random tune algorithm
random.tune <- makeTuneControlRandom(maxit = 5)
```

```{r}
set.seed(1)

#Defining learner model
lrn2 <- makeLearner("regr.rpart") 

#Get parameter set of the learner
getParamSet(lrn2)                        

# Define parameters of model 
model.params <- makeParamSet( 
  makeNumericParam("cp", lower = 0, upper = 0.1),
  makeIntegerParam("minsplit", lower = 5, upper = 20),
  makeIntegerParam("minbucket", lower = 2, upper = 10)
)

#Hyperparameter tuning
tuned.model2 <- tuneParams(learner    = lrn2,
                             task       = regr.task.train2,
                             resampling = cv.folds,
                             par.set    = model.params,
                             control    = random.tune,
                             measures = list(rmse,mae),
                             show.info  = TRUE)
tuned.model2$x
```
############# Evaluation of Model #############
```{r}

lrn2 <- makeLearner("regr.rpart",par.vals = tuned.model2$x)

res2 <- mlr::train(lrn2,regr.task.train2)
rpc2 = rpart.control(minsplit = tuned.model2$x$minsplit, minbucket = tuned.model2$x$minbucket, cp = tuned.model2$x$cp
                    )
mod2 <- rpart(Max.Temperature ~ ., data = train2, control = rpc2)

#Calucating the RMSE of training set 
rmse_reg(mod2, testing = train2, target = "Max.Temperature")
getLearnerModel(res2)

rpart.plot(res2$learner.model,roundint=FALSE)

```

```{r}

(mlrpred2   <- predict(res2, newdata = test2))

#Calcuating the RMSE of test set

sc2 <- performance(mlrpred2, measures = rmse) 

print(sc2)

```

############### Station code: Godsamba ###############

```{r}
my_data3 <- read_excel('SCGod.xlsx')
my_data3 <- as.data.frame(my_data3)

glimpse(my_data3)

my_data3$Station.Code<-NULL
my_data3$Hour<-NULL
my_data3$Year<-NULL
my_data3$Month<-NULL
my_data3$Day<-NULL

str(my_data3)
```

```{r}
# change categorical values to numerical values

my_data3[my_data3=="C"]<-0
my_data3[my_data3=="E"]<-9
my_data3[my_data3=="S"]<-18
my_data3[my_data3=="W"]<-27
my_data3[my_data3=="N"]<-36
my_data3[my_data3=="NNE"]<-2
my_data3[my_data3=="ESE"]<-11
my_data3[my_data3=="SSW"]<- 20
my_data3[my_data3=="WNW"]<- 29
my_data3[my_data3=="V"]<- 99
my_data3[my_data3=="NE"]<- 5
my_data3[my_data3=="SE"]<- 14
my_data3[my_data3=="SW"]<- 23
my_data3[my_data3=="NW"]<- 32
my_data3[my_data3=="ENE"]<- 7
my_data3[my_data3=="SSE"]<- 16
my_data3[my_data3=="WSW"]<- 25
my_data3[my_data3=="NNW"]<- 34

my_data3$Wind.direction <- as.numeric(as.character(my_data3$Wind.direction))

str(my_data3)
```
################## Histograms distribution of data ##################
```{r}
ggplot(my_data3, aes(x=my_data3$Min.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data3, aes(x=my_data3$Max.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data3, aes(x=my_data3$Temp.Dry.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data3, aes(x=my_data3$Temp.Wet.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data3, aes(x=my_data3$Inst.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data3, aes(x=my_data3$Av.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data3, aes(x=my_data3$Relative.Humidity)) + geom_histogram()

```
```{r}
ggplot(my_data3, aes(x=my_data3$Temp...Pan.Water)) + geom_histogram()

```
```{r}
ggplot(my_data3, aes(x=my_data3$Pan.Evaporation)) + geom_histogram()

```
################ Visualizing the Correlation matrix ################
```{r}
source("http://www.sthda.com/upload/rquery_cormat.r")
require("corrplot")
rquery.cormat(my_data3)
```

```{r}
#Splitting the dataset

sample3 <- sample.int(n = nrow(my_data3), size = floor(.75*nrow(my_data3)), replace = F)
train3 <- my_data3[sample3, ]
test3  <- my_data3[-sample3, ]

#Regression Tree:Task Creation
regr.task.train3<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(train3), 
                        target = "Max.Temperature")

regr.task.test3<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(test3), 
                        target = "Max.Temperature")

#Repeated cross validation
cv.folds <- makeResampleDesc("CV", iters = 3)

# Define model tuning algorithm ~ Random tune algorithm
random.tune <- makeTuneControlRandom(maxit = 5)

```

```{r}
set.seed(1)

#Defining learner model
lrn3 <- makeLearner("regr.rpart") 

#Get parameter set of the learner
getParamSet(lrn3)                        

# Define parameters of model  
model.params <- makeParamSet( 
  makeNumericParam("cp", lower = 0, upper = 0.1),
  makeIntegerParam("minsplit", lower = 5, upper = 20),
  makeIntegerParam("minbucket", lower = 2, upper = 10)
)

#Hyperparameter tuning
tuned.model3 <- tuneParams(learner    = lrn3,
                             task       = regr.task.train3,
                             resampling = cv.folds,
                             par.set    = model.params,
                             control    = random.tune,
                             measures = list(rmse,mae),
                             show.info  = TRUE)
tuned.model3$x
```
############# Evaluation of Model #############
```{r}
lrn3 <- makeLearner("regr.rpart",par.vals = tuned.model3$x)

res3 <- mlr::train(lrn3,regr.task.train3)
rpc3 = rpart.control(minsplit = tuned.model3$x$minsplit, minbucket = tuned.model3$x$minbucket, cp = tuned.model3$x$cp
                    )
mod3 <- rpart(Max.Temperature ~ ., data = train3, control = rpc3)

#Calucating the RMSE of training set 
rmse_reg(mod3, testing = train3, target = "Max.Temperature")
getLearnerModel(res3)

rpart.plot(res3$learner.model,roundint=FALSE)

```

```{r}

(mlrpred3   <- predict(res3, newdata = test3))

#Calcuating the RMSE of test set

sc3<- performance(mlrpred1, measures = rmse) 

print(sc3)

```

############### Station code: Kakrapar ###############

```{r}
my_data4 <- read_excel('SCKakra.xlsx')
my_data4 <- as.data.frame(my_data4)

glimpse(my_data4)

my_data4$Station.Code<-NULL
my_data4$Hour<-NULL
my_data4$Year<-NULL
my_data4$Month<-NULL
my_data4$Day<-NULL

str(my_data4)
```

```{r}
# change categorical values to numerical values

my_data4[my_data4=="C"]<-0
my_data4[my_data4=="E"]<-9
my_data4[my_data4=="S"]<-18
my_data4[my_data4=="W"]<-27
my_data4[my_data4=="N"]<-36
my_data4[my_data4=="NNE"]<-2
my_data4[my_data4=="ESE"]<-11
my_data4[my_data4=="SSW"]<- 20
my_data4[my_data4=="WNW"]<- 29
my_data4[my_data4=="V"]<- 99
my_data4[my_data4=="NE"]<- 5
my_data4[my_data4=="SE"]<- 14
my_data4[my_data4=="SW"]<- 23
my_data4[my_data4=="NW"]<- 32
my_data4[my_data4=="ENE"]<- 7
my_data4[my_data4=="SSE"]<- 16
my_data4[my_data4=="WSW"]<- 25
my_data4[my_data4=="NNW"]<- 34

my_data4$Wind.direction <- as.numeric(as.character(my_data4$Wind.direction))

str(my_data4)
```
################## Histograms distribution of data ##################
```{r}
ggplot(my_data4, aes(x=my_data4$Min.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data4, aes(x=my_data4$Max.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data4, aes(x=my_data4$Temp.Dry.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data4, aes(x=my_data4$Temp.Wet.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data4, aes(x=my_data4$Inst.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data4, aes(x=my_data4$Av.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data4, aes(x=my_data4$Relative.Humidity)) + geom_histogram()

```
```{r}
ggplot(my_data4, aes(x=my_data4$Temp...Pan.Water)) + geom_histogram()

```
```{r}
ggplot(my_data4, aes(x=my_data4$Pan.Evaporation)) + geom_histogram()

```
################ Visualizing the Correlation matrix ################
```{r}
source("http://www.sthda.com/upload/rquery_cormat.r")
require("corrplot")
rquery.cormat(my_data4)
```

```{r}
#Splitting the dataset

sample4 <- sample.int(n = nrow(my_data4), size = floor(.75*nrow(my_data4)), replace = F)
train4 <- my_data4[sample4, ]
test4  <- my_data4[-sample4, ]

#Regression Tree:Task Creation
regr.task.train4<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(train4), 
                        target = "Max.Temperature")

regr.task.test4<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(test4), 
                        target = "Max.Temperature")

#Repeated cross validation
cv.folds <- makeResampleDesc("CV", iters = 3)

# Define model tuning algorithm ~ Random tune algorithm
random.tune <- makeTuneControlRandom(maxit = 5)
```

```{r}
set.seed(1)

#Defining learner model
lrn4 <- makeLearner("regr.rpart") 

#Get parameter set of the learner
getParamSet(lrn4)                        

# Define parameters of model and search grid 
model.params <- makeParamSet( 
  makeNumericParam("cp", lower = 0, upper = 0.1),
  makeIntegerParam("minsplit", lower = 5, upper = 20),
  makeIntegerParam("minbucket", lower = 2, upper = 10)
)

#Hyperparameter tuning
tuned.model4 <- tuneParams(learner    = lrn4,
                             task       = regr.task.train4,
                             resampling = cv.folds,
                             par.set    = model.params,
                             control    = random.tune,
                             measures = list(rmse,mae),
                             show.info  = TRUE)
tuned.model4$x
```
############# Evaluation of Model #############
```{r}
lrn4 <- makeLearner("regr.rpart",par.vals = tuned.model4$x)

res4 <- mlr::train(lrn4,regr.task.train4)
rpc4 = rpart.control(minsplit = tuned.model4$x$minsplit, minbucket = tuned.model4$x$minbucket, cp = tuned.model4$x$cp
                    )
mod4 <- rpart(Max.Temperature ~ ., data = train4, control = rpc4)

#Calucating the RMSE of training set 
rmse_reg(mod4, testing = train4, target = "Max.Temperature")
getLearnerModel(res3)

rpart.plot(res4$learner.model,roundint=FALSE)

```

```{r}


(mlrpred4   <- predict(res4, newdata = test4))

#Calcuating the RMSE of test set

sc4 <- performance(mlrpred4, measures = rmse)

print(sc4)

```

############### Station code: Rander ###############

```{r}
my_data5 <- read_excel('SCRan.xlsx')
my_data5 <- as.data.frame(my_data5)

glimpse(my_data5)

my_data5$Station.Code<-NULL
my_data5$Hour<-NULL
my_data5$Year<-NULL
my_data5$Month<-NULL
my_data5$Day<-NULL

str(my_data5)
```

```{r}
# change categorical values to numerical values

my_data5[my_data5=="C"]<-0
my_data5[my_data5=="E"]<-9
my_data5[my_data5=="S"]<-18
my_data5[my_data5=="W"]<-27
my_data5[my_data5=="N"]<-36
my_data5[my_data5=="NNE"]<-2
my_data5[my_data5=="ESE"]<-11
my_data5[my_data5=="SSW"]<- 20
my_data5[my_data5=="WNW"]<- 29
my_data5[my_data5=="V"]<- 99
my_data5[my_data5=="NE"]<- 5
my_data5[my_data5=="SE"]<- 14
my_data5[my_data5=="SW"]<- 23
my_data5[my_data5=="NW"]<- 32
my_data5[my_data5=="ENE"]<- 7
my_data5[my_data5=="SSE"]<- 16
my_data5[my_data5=="WSW"]<- 25
my_data5[my_data5=="NNW"]<- 34

my_data5$Wind.direction <- as.numeric(as.character(my_data5$Wind.direction))

str(my_data5)
```
################## Histograms distribution of data ##################
```{r}
ggplot(my_data5, aes(x=my_data5$Min.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data5, aes(x=my_data5$Max.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data5, aes(x=my_data5$Temp.Dry.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data5, aes(x=my_data5$Temp.Wet.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data5, aes(x=my_data5$Inst.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data5, aes(x=my_data5$Av.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data5, aes(x=my_data5$Relative.Humidity)) + geom_histogram()

```
```{r}
ggplot(my_data5, aes(x=my_data5$Temp...Pan.Water)) + geom_histogram()

```
```{r}
ggplot(my_data5, aes(x=my_data5$Pan.Evaporation)) + geom_histogram()

```
################ Visualizing the Correlation matrix ################
```{r}
source("http://www.sthda.com/upload/rquery_cormat.r")
require("corrplot")
rquery.cormat(my_data5)
```

```{r}
#Splitting the dataset

sample5 <- sample.int(n = nrow(my_data5), size = floor(.75*nrow(my_data5)), replace = F)
train5 <- my_data5[sample5, ]
test5  <- my_data5[-sample5, ]

#Regression Tree:Task Creation
regr.task.train5<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(train5), 
                        target = "Max.Temperature")

regr.task.test5<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(test5), 
                        target = "Max.Temperature")

#Repeated cross validation
cv.folds <- makeResampleDesc("CV", iters = 3)

# Define model tuning algorithm ~ Random tune algorithm
random.tune <- makeTuneControlRandom(maxit = 5)
```

```{r}
set.seed(1)

#Defining learner model
lrn5 <- makeLearner("regr.rpart") 

#Get parameter set of the learner
getParamSet(lrn4)                        

# Define parameters of model and search grid 
model.params <- makeParamSet( 
  makeNumericParam("cp", lower = 0, upper = 0.1),
  makeIntegerParam("minsplit", lower = 5, upper = 20),
  makeIntegerParam("minbucket", lower = 2, upper = 10)
)

#Hyperparameter tuning
tuned.model5 <- tuneParams(learner    = lrn5,
                             task       = regr.task.train5,
                             resampling = cv.folds,
                             par.set    = model.params,
                             control    = random.tune,
                             measures = list(rmse,mae),
                             show.info  = TRUE)
tuned.model5$x
```
############# Evaluation of Model #############
```{r}
lrn5 <- makeLearner("regr.rpart",par.vals = tuned.model5$x)

res5 <- mlr::train(lrn5,regr.task.train5)
rpc5 = rpart.control(minsplit = tuned.model5$x$minsplit, minbucket = tuned.model5$x$minbucket, cp = tuned.model5$x$cp
                    )
mod5 <- rpart(Max.Temperature ~ ., data = train5, control = rpc5)

#Calucating the RMSE of training set 
rmse_reg(mod5, testing = train5, target = "Max.Temperature")
getLearnerModel(res5)

rpart.plot(res5$learner.model,roundint=FALSE)

```

```{r}

(mlrpred5   <- predict(res5, newdata = test5))

#Calcuating the RMSE of test set

sc5 <- performance(mlrpred5, measures = rmse)

print(sc5)

```

############### Station code: Ukai ###############

```{r}
my_data6 <- read_excel('SCUkai.xlsx')
my_data6 <- as.data.frame(my_data6)

glimpse(my_data6)

my_data6$Station.Code<-NULL
my_data6$Hour<-NULL
my_data6$Year<-NULL
my_data6$Month<-NULL
my_data6$Day<-NULL

str(my_data6)
```

```{r}
# change categorical values to numerical values

my_data6[my_data6=="C"]<-0
my_data6[my_data6=="E"]<-9
my_data6[my_data6=="S"]<-18
my_data6[my_data6=="W"]<-27
my_data6[my_data6=="N"]<-36
my_data6[my_data6=="NNE"]<-2
my_data6[my_data6=="ESE"]<-11
my_data6[my_data6=="SSW"]<- 20
my_data6[my_data6=="WNW"]<- 29
my_data6[my_data6=="V"]<- 99
my_data6[my_data6=="NE"]<- 5
my_data6[my_data6=="SE"]<- 14
my_data6[my_data6=="SW"]<- 23
my_data6[my_data6=="NW"]<- 32
my_data6[my_data6=="ENE"]<- 7
my_data6[my_data6=="SSE"]<- 16
my_data6[my_data6=="WSW"]<- 25
my_data6[my_data6=="NNW"]<- 34

my_data6$Wind.direction <- as.numeric(as.character(my_data6$Wind.direction))

str(my_data6)
```
################## Histograms distribution of data ##################
```{r}
ggplot(my_data6, aes(x=my_data6$Min.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data6, aes(x=my_data6$Max.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data6, aes(x=my_data6$Temp.Dry.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data6, aes(x=my_data6$Temp.Wet.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data6, aes(x=my_data6$Inst.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data6, aes(x=my_data6$Av.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data6, aes(x=my_data6$Relative.Humidity)) + geom_histogram()

```
```{r}
ggplot(my_data6, aes(x=my_data6$Temp...Pan.Water)) + geom_histogram()

```
```{r}
ggplot(my_data6, aes(x=my_data6$Pan.Evaporation)) + geom_histogram()

```
################ Visualizing the Correlation matrix ################
```{r}
source("http://www.sthda.com/upload/rquery_cormat.r")
require("corrplot")
rquery.cormat(my_data6)
```

```{r}
#Splitting the dataset

sample6 <- sample.int(n = nrow(my_data6), size = floor(.75*nrow(my_data6)), replace = F)
train6 <- my_data6[sample6, ]
test6  <- my_data6[-sample6, ]

#Regression Tree:Task Creation
regr.task.train6<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(train6), 
                        target = "Max.Temperature")

regr.task.test6<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(test6), 
                        target = "Max.Temperature")

#Repeated cross validation
cv.folds <- makeResampleDesc("CV", iters = 3)

# Define model tuning algorithm ~ Random tune algorithm
random.tune <- makeTuneControlRandom(maxit = 5)
```

```{r}
set.seed(1)

#Defining learner model
lrn6 <- makeLearner("regr.rpart") 

#Get parameter set of the learner
getParamSet(lrn6)                        

# Define parameters of model
model.params <- makeParamSet( 
  makeNumericParam("cp", lower = 0, upper = 0.1),
  makeIntegerParam("minsplit", lower = 5, upper = 20),
  makeIntegerParam("minbucket", lower = 2, upper = 10)
)

#Hyperparameter tuning
tuned.model6 <- tuneParams(learner    = lrn6,
                             task       = regr.task.train6,
                             resampling = cv.folds,
                             par.set    = model.params,
                             control    = random.tune,
                             measures = list(rmse,mae),
                             show.info  = TRUE)
tuned.model6$x
```
############# Evaluation of Model #############
```{r}
lrn6 <- makeLearner("regr.rpart",par.vals = tuned.model6$x)

res6 <- mlr::train(lrn6,regr.task.train6)
rpc6 = rpart.control(minsplit = tuned.model6$x$minsplit, minbucket = tuned.model6$x$minbucket, cp = tuned.model6$x$cp
                    )
mod6 <- rpart(Max.Temperature ~ ., data = train6, control = rpc6)

#Calucating the RMSE of training set
rmse_reg(mod6, testing = train6, target = "Max.Temperature")
getLearnerModel(res6)

rpart.plot(res6$learner.model,roundint=FALSE)

```

```{r}

(mlrpred6   <- predict(res6, newdata = test6))

#Calcuating the RMSE of test set
sc6 <- performance(mlrpred6, measures = rmse)
print(sc6)

```

############### Station code: Olpad ###############

```{r}
my_data7 <- read_excel('SCOlpad.xlsx')
my_data7 <- as.data.frame(my_data7)

glimpse(my_data7)

my_data7$Station.Code<-NULL
my_data7$Hour<-NULL
my_data7$Year<-NULL
my_data7$Month<-NULL
my_data7$Day<-NULL

str(my_data7)
```

```{r}
# change categorical values to numerical values

my_data7[my_data7=="C"]<-0
my_data7[my_data7=="E"]<-9
my_data7[my_data7=="S"]<-18
my_data7[my_data7=="W"]<-27
my_data7[my_data7=="N"]<-36
my_data7[my_data7=="NNE"]<-2
my_data7[my_data7=="ESE"]<-11
my_data7[my_data7=="SSW"]<- 20
my_data7[my_data7=="WNW"]<- 29
my_data7[my_data7=="V"]<- 99
my_data7[my_data7=="NE"]<- 5
my_data7[my_data7=="SE"]<- 14
my_data7[my_data7=="SW"]<- 23
my_data7[my_data7=="NW"]<- 32
my_data7[my_data7=="ENE"]<- 7
my_data7[my_data7=="SSE"]<- 16
my_data7[my_data7=="WSW"]<- 25
my_data7[my_data7=="NNW"]<- 34

my_data7$Wind.direction <- as.numeric(as.character(my_data7$Wind.direction))

str(my_data7)
```
################## Histograms distribution of data ##################
```{r}
ggplot(my_data7, aes(x=my_data7$Min.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data7, aes(x=my_data7$Max.Temperature)) + geom_histogram()

```
```{r}
ggplot(my_data7, aes(x=my_data7$Temp.Dry.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data7, aes(x=my_data7$Temp.Wet.Bulb)) + geom_histogram()

```
```{r}
ggplot(my_data7, aes(x=my_data7$Inst.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data7, aes(x=my_data7$Av.Wind.Speed)) + geom_histogram()

```
```{r}
ggplot(my_data7, aes(x=my_data7$Relative.Humidity)) + geom_histogram()

```
```{r}
ggplot(my_data7, aes(x=my_data7$Temp...Pan.Water)) + geom_histogram()

```
```{r}
ggplot(my_data7, aes(x=my_data7$Pan.Evaporation)) + geom_histogram()

```
################ Visualizing the Correlation matrix ################
```{r}
source("http://www.sthda.com/upload/rquery_cormat.r")
require("corrplot")
rquery.cormat(my_data7)
```

```{r}
#Splitting the dataset

sample7 <- sample.int(n = nrow(my_data7), size = floor(.75*nrow(my_data7)), replace = F)
train7 <- my_data7[sample7, ]
test7  <- my_data7[-sample7, ]

#Regression Tree:Task Creation
regr.task.train7<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(train7), 
                        target = "Max.Temperature")

regr.task.test7<-makeRegrTask(id  = "Tin_regr", 
                        data   = data.frame(test7), 
                        target = "Max.Temperature")

#Repeated cross validation
cv.folds <- makeResampleDesc("CV", iters = 3)

# Define model tuning algorithm ~ Random tune algorithm
random.tune <- makeTuneControlRandom(maxit = 5)
```

```{r}
set.seed(1)

#Defining learner model
lrn7 <- makeLearner("regr.rpart") 

#Get parameter set of the learner
getParamSet(lrn7)                        

# Define parameters of model 
model.params <- makeParamSet( 
  makeNumericParam("cp", lower = 0, upper = 0.1),
  makeIntegerParam("minsplit", lower = 5, upper = 20),
  makeIntegerParam("minbucket", lower = 2, upper = 10)
)

#Hyperparameter tuning
tuned.model7 <- tuneParams(learner    = lrn7,
                             task       = regr.task.train7,
                             resampling = cv.folds,
                             par.set    = model.params,
                             control    = random.tune,
                             measures = list(rmse,mae),
                             show.info  = TRUE)
tuned.model7$x
```
############# Evaluation of Model #############
```{r}

lrn7 <- makeLearner("regr.rpart",par.vals = tuned.model7$x)

res7 <- mlr::train(lrn7,regr.task.train7)
rpc7 = rpart.control(minsplit = tuned.model7$x$minsplit, minbucket = tuned.model7$x$minbucket, cp = tuned.model7$x$cp
                    )
mod7 <- rpart(Max.Temperature ~ ., data = train7, control = rpc7)

#Calucating the RMSE of training set 
rmse_reg(mod7, testing = train7, target = "Max.Temperature")
getLearnerModel(res7)

rpart.plot(res7$learner.model,roundint=FALSE)

```

```{r}

(mlrpred7 <- predict(res7, newdata = test7))

#Calcuating the RMSE of test set
sc7 <- performance(mlrpred7, measures = rmse)

print(sc7)

```
############# Cummulative RMSE of all Station Code #############
```{r}

RMSE = (sc1+sc2+sc3+sc4+sc5+sc6+sc7)/7
print(RMSE)

```

