---
title: "TrendAnalysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Install and load required package libraries
```{r}
#install.packages("Kendall")
#install.packages("trend")

library(Kendall)
library(modifiedmk)
library(trend)
library(tidyverse)
library(dplyr)
```

# Read data for trend analysis
```{r}
library("readxl")
trend_analysis <- read_excel("FinalModifiedData.xlsx")
tadata = data.frame(trend_analysis)

print(tadata)
```
# Select Distinct Station codes from Dataframe
```{r}
tadata %>% distinct(Station.Code)
```

# Station wise data
```{r}
Amli_data <- filter(tadata, Station.Code == 'Amli')
Chopadvav_data	<- filter(tadata, Station.Code == 'Chopadvav')			
Godsamba_data <- filter(tadata, Station.Code == 'Godsamba')				
Kakrapar_data <- filter(tadata, Station.Code == 'Kakrapar')				
Olpad_data	<- filter(tadata, Station.Code == 'Olpad')			
Rander_data <- filter(tadata, Station.Code == 'Rander')				
Ukai_data <- filter(tadata, Station.Code == 'Ukai')
print(Amli_data)
print(Chopadvav_data)
print(Godsamba_data)
print(Kakrapar_data)
print(Olpad_data)
print(Rander_data)
print(Ukai_data)
```

# Station wise yearly trend analysis of max temp
# For Station Code "AMLI"
```{r}
df_Amli <- Amli_data %>% 
            group_by(Year,Month) %>% 
              summarize(Mean_MaxTemp = mean(Max.Temperature, na.rm = TRUE))

print(df_Amli)

TS_Amli = ts(df_Amli$Mean_MaxTemp,
        frequency=12,
        start=c(2000))

print(TS_Amli)
```

# Plot the graph
```{r}
plot(TS_Amli)
lines(lowess(time(TS_Amli),TS_Amli), col='blue')
plot(decompose(TS_Amli))
plot(stl(TS_Amli, s.window="periodic"))
```

# Trend Analysis 
```{r}
###  tau is a measure of the strength of the trend
###  p-value for trend is also shown
### A very high positive value of S is an indicator of an increasing trend and a very low negative value indicates a decreasing trend. The presence of a statistically significant trend is evaluated using Z value.
###  tau is a measure of the strength of the trend
###  p-value for trend is also shown

print('Mann- Kendall Test for Amli')
MK_Amli = MannKendall(TS_Amli)
summary(MK_Amli)

mk.test(TS_Amli, continuity = TRUE)

cat('\n')
print('Seasonal Mann- Kendall Test for Amli')
SMK_Amli = SeasonalMannKendall(TS_Amli)
summary(SMK_Amli)

# print('Vector Value')
Amli_vec <- df_Amli[['Mean_MaxTemp']]
# mkttest(Amli_vec)

### Modified Mann-Kendall Test For Serially Correlated Data Using the Hamed and Rao (1998) Variance Correction Approach
cat('\n')
print('Modified MK Test')
mmkh(Amli_vec,ci=0.95)

### Mann-Kendall Test of Prewhitened Time Series Data in Presence of Serial Correlation Using the von Storch (1995) Approach
cat('\n')
print('Modified MK Test with prewhitening')
pwmk(Amli_vec)

cat('\n')
print('Spearman Coefficient')
spear(Amli_vec)

sens.slope(TS_Amli)
### A positive value of Sen's slope indicates an upward or increasing trend and a negative value gives a downward or decreasing trend in the time series.

pettitt.test(TS_Amli)

```

# Point of Change for Station Amli
```{r}
df_Amli[66,]
```

# For Station Code "CHOPADVAV"
```{r}
df_Chop <- Chopadvav_data %>% 
            group_by(Year,Month) %>% 
              summarize(Mean_MaxTemp = mean(Max.Temperature, na.rm = TRUE))

print(df_Chop)

TS_Chop = ts(df_Chop$Mean_MaxTemp,
        frequency=12,
        start=c(1999))

print(TS_Chop)
```

# Plot the graph
```{r}
plot(TS_Chop)
lines(lowess(time(TS_Chop),TS_Chop), col='blue')
plot(decompose(TS_Chop))
plot(stl(TS_Chop, s.window="periodic"))
```

# Trend Analysis 
```{r}
print('Mann- Kendall Test for Chopadvav')
MK_Chop = MannKendall(TS_Chop)
summary(MK_Chop)

mk.test(TS_Chop, continuity = TRUE)

print('Seasonal Mann- Kendall Test for Chopadvav')
SMK_Chop = SeasonalMannKendall(TS_Chop)
summary(SMK_Chop)

Chop_vec <- df_Chop[['Mean_MaxTemp']]

cat('\n')
print('Modified MK Test')
mmkh(Chop_vec,ci=0.95)

cat('\n')
print('Modified MK Test with prewhitening')
pwmk(Chop_vec)

sens.slope(TS_Chop)

pettitt.test(TS_Chop)

```
# Point of Change for Station Chopadvav
```{r}
df_Chop[54,]
```

# For Station Code "GODSAMBA"				
```{r}
df_God <- Godsamba_data %>% 
            group_by(Year,Month) %>% 
              summarize(Mean_MaxTemp = mean(Max.Temperature, na.rm = TRUE))

print(df_God)

TS_God = ts(df_God$Mean_MaxTemp,
        frequency=12,
        start=c(1999))

print(TS_God)
```

# Plot the graph
```{r}
plot(TS_God)
lines(lowess(time(TS_God),TS_God), col='blue')
plot(decompose(TS_God))
plot(stl(TS_God, s.window="periodic"))
```

# Trend Analysis 
```{r}
print('Mann- Kendall Test for Godsamba')
MK_God = MannKendall(TS_God)
summary(MK_God)

mk.test(TS_God, continuity = TRUE)

print('Seasonal Mann- Kendall Test for Godsamba')
SMK_God = SeasonalMannKendall(TS_Chop)
summary(SMK_God)

God_vec <- df_God[['Mean_MaxTemp']]

cat('\n')
print('Modified MK Test')
mmkh(God_vec,ci=0.95)

cat('\n')
print('Modified MK Test with prewhitening')
pwmk(God_vec)

sens.slope(TS_God)

pettitt.test(TS_God)
```
# Point of Change for Station Godsamba
```{r}
df_God[24,]
```

# For Station Code "KAKRAPAR"				

```{r}
df_Kak <- Kakrapar_data %>% 
            group_by(Year,Month) %>% 
              summarize(Mean_MaxTemp = mean(Max.Temperature, na.rm = TRUE))

print(df_Kak)

TS_Kak = ts(df_Kak$Mean_MaxTemp,
        frequency=12,
        start=c(2000))

print(TS_Kak)
```

# Plot the graph
```{r}
plot(TS_Kak)
lines(lowess(time(TS_Kak),TS_Kak), col='blue')
plot(decompose(TS_Kak))
plot(stl(TS_Kak, s.window="periodic"))
```

# Trend Analysis 
```{r}
print('Mann- Kendall Test for Kakrapar')
MK_Kak = MannKendall(TS_Kak)
summary(MK_Kak)

mk.test(TS_Kak, continuity = TRUE)

print('Seasonal Mann- Kendall Test for Kakrapar')
SMK_Kak = SeasonalMannKendall(TS_Kak)
summary(SMK_Kak)

Kak_vec <- df_Kak[['Mean_MaxTemp']]

cat('\n')
print('Modified MK Test')
mmkh(Kak_vec,ci=0.95)

cat('\n')
print('Modified MK Test with prewhitening')
pwmk(Kak_vec)

sens.slope(TS_Kak)

pettitt.test(TS_Kak)
```

# Point of Change for Station Kakrapar
```{r}
df_Kak[81,]
```

# For Station Code "OLPAD"				

```{r}
df_Olp <- Olpad_data %>% 
            group_by(Year,Month) %>% 
              summarize(Mean_MaxTemp = mean(Max.Temperature, na.rm = TRUE))

print(df_Olp)

TS_Olp = ts(df_Olp$Mean_MaxTemp,
        frequency=12,
        start=c(1999))

print(TS_Olp)
```

# Plot the graph
```{r}
plot(TS_Olp)
lines(lowess(time(TS_Olp),TS_Olp), col='blue')
plot(decompose(TS_Olp))
plot(stl(TS_Olp, s.window="periodic"))
```

# Trend Analysis 
```{r}
print('Mann- Kendall Test for Olpad')
MK_Olp = MannKendall(TS_Olp)
summary(MK_Olp)

mk.test(TS_Chop, continuity = TRUE)

print('Seasonal Mann- Kendall Test for Olpad')
SMK_Olp = SeasonalMannKendall(TS_Olp)
summary(SMK_Olp)

Olp_vec <- df_Olp[['Mean_MaxTemp']]

cat('\n')
print('Modified MK Test')
mmkh(Olp_vec,ci=0.95)

cat('\n')
print('Modified MK Test with prewhitening')
pwmk(Olp_vec)

sens.slope(TS_Olp)

pettitt.test(TS_Olp)
```

# Point of Change for Station Olpad
```{r}
df_Olp[33,]
```

# For Station Code "RANDER"			

```{r}
df_Ran <- Rander_data %>% 
            group_by(Year,Month) %>% 
              summarize(Mean_MaxTemp = mean(Max.Temperature, na.rm = TRUE))

print(df_Ran)

TS_Ran = ts(df_Ran$Mean_MaxTemp,
        frequency=12,
        start=c(1999))

print(TS_Ran)
```

# Plot the graph
```{r}
plot(TS_Ran)
lines(lowess(time(TS_Ran),TS_Ran), col='blue')
plot(decompose(TS_Ran))
plot(stl(TS_Ran, s.window="periodic"))
```

# Trend Analysis 
```{r}
print('Mann- Kendall Test for Rander')
MK_Ran = MannKendall(TS_Ran)
summary(MK_Ran)

mk.test(TS_Ran, continuity = TRUE)

print('Seasonal Mann- Kendall Test for Rander')
SMK_Ran = SeasonalMannKendall(TS_Ran)
summary(SMK_Ran)

Ran_vec <- df_Ran[['Mean_MaxTemp']]

cat('\n')
print('Modified MK Test')
mmkh(Ran_vec,ci=0.95)

cat('\n')
print('Modified MK Test with prewhitening')
pwmk(Ran_vec)

sens.slope(TS_Ran)

pettitt.test(TS_Ran)
```

# Point of Change for Station Rander
```{r}
df_Ran[46,]
```

# For Station Code "Ukai"				
```{r}
df_Ukai <- Ukai_data %>% 
            group_by(Year,Month) %>% 
              summarize(Mean_MaxTemp = mean(Max.Temperature, na.rm = TRUE))

print(df_Ukai)

TS_Ukai = ts(df_Ukai$Mean_MaxTemp,
        frequency=12,
        start=c(1999))

print(TS_Ukai)
```

# Plot the graph
```{r}
plot(TS_Ukai)
lines(lowess(time(TS_Ukai),TS_Ukai), col='blue')
plot(decompose(TS_Ukai))
plot(stl(TS_Ukai, s.window="periodic"))
```

# Trend Analysis 
```{r}
print('Mann- Kendall Test for Ukai')
MK_Ukai = MannKendall(TS_Ukai)
summary(MK_Ukai)

mk.test(TS_Ukai, continuity = TRUE)

print('Seasonal Mann- Kendall Test for Ukai')
SMK_Ukai = SeasonalMannKendall(TS_Ukai)
summary(SMK_Ukai)

Ukai_vec <- df_Ukai[['Mean_MaxTemp']]

cat('\n')
print('Modified MK Test')
mmkh(Ukai_vec,ci=0.95)

cat('\n')
print('Modified MK Test with prewhitening')
pwmk(Ukai_vec)

sens.slope(TS_Ukai)

pettitt.test(TS_Ukai)
```

# Point of Change for Station Ukai
```{r}
df_Ukai[50,]
```












